---
title: "Pivot Lobsters"
author: "Julia Large"
date: "1/27/2020"
output: html_document
---

```{r}

library(here)
library(tidyverse)
library(readxl)
library(skimr)
library(kableExtra)

lobsters <- read_xlsx(here("data", "lobsters2.xlsx"),
                      skip = 4) # skip the first 4 rows of the file

```

### Viewing our data with skimr

```{r}

skimr::skim(lobsters) #the "::" notation makes it clear which package the function is being pulled from, which is useful when you're using a function which you don't commonly use

#the above is the same as the below:

skim(lobsters)

```


### Pivot Tables
In R, pivot tables are: group_by() %>% summarize()
Shortcut for pipes: CTRL-shift-m
'n()' function counts number of variables/observations

```{r}

#pivot table grouped by year
lobsters %>% 
  group_by(year) %>% 
  summarize(count_by_year = n())

#pivot table grouped by year and site
lobsters %>% 
  group_by(site,
           year) %>% 
  summarize(count_by_year_site = n())

```

### Summarize by multiple variables and save as an object (siteyear_summary)

To exclude missing values which would prevent the calculation from happening, use na.rm = TRUE 
(in excel, this would be iferror() )
(na.rm stands for "non-applicable.remove")

```{r}

siteyear_summary <- lobsters %>% 
  group_by(site,
           year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))

```

### Table formatting with `kableExtra::kable()`

```{r}

siteyear_summary %>% 
  kableExtra::kable()

```

#### There are `r nrow(lobsters)` total lobsters included in this report.

```{r}

med_lobsters <- lobsters %>% 
  group_by(site,
           year) %>% 
  summarize(med_size_mm = median(size_mm, na.rm = TRUE))

ggplot(med_lobsters, aes(year, med_size_mm)) + 
  geom_point(aes(color = site,
                 size = med_size_mm))

```







