---
title: "Creating graphs with ggplot2"
author: "Julia Large"
date: "1/27/2020"
output: html_document
---

### Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.

We will install these packages:

- tidyverse
- readxl
- here

```{r setup, message=FALSE}

library(tidyverse) #install.packages("tidyverse")
library(readxl) #install.packages("readxl")
library(here) #install.packages("here")

```

### California National Parks Data

```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
#also can be written as:
#ca_np <- read_csv("data/ca_np.csv")

ci_np <- read_excel(here("data/ci_np.xlsx"))

```


### Make my own first graph!

We're going to make a scatterplot of year (x-var) versus visitors (y-var) in ggplot:

Note: Command + Alt + I = insert new code chunk

```{r}

ggplot(data = ci_np, mapping = aes(year, visitors)) +
  geom_point()


```

### Creating the base layer of my graphs as an object
which makes it easier to use for different types of graphs

```{r}

gg_base <- ggplot(data = ci_np, mapping = aes(year, visitors))

gg_base + 
  geom_point()

```

### Now let's start customizing!

Type of argument to use:
- Line/point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha
- Change size: size

```{r}

gg_base + 
  geom_point(
    color = "blue",
    linetype = "dashed",
    pch = 8,
    size = 3
  )

```

### Update graph aesthetics based on variables

To do: wrap the argumnet in "aes()"

```{r}

gg_base +
  geom_point(
    aes(size = visitors,
        color = year),
    pch = 8
  )

```

### Use gg-themes for overall updates

- theme_light()
- theme_minimal()
- theme_bw() : ie, black & white

```{r}

gg_base + 
  geom_line(
    aes(size = visitors,
        color = year),
    pch = 8
  ) + 
  theme_linedraw()

```

### Create a multi-series line plot using data for all CA National Parks

```{r}

ggplot(ca_np, aes(year, visitors, group = park_name)) + 
  geom_line(
    aes(color = park_name,
        size = visitors)
  )

```

### Create a different graph for each type of one variable

Use facet_wrap(~) which creates a separate chart for each type of "facet"

Remove the legends by : show.legend = FALSE

```{r}

ggplot(ca_np, aes(year, visitors, group = park_name)) + 
  geom_point(
    aes(color = park_name,
        size = visitors),
    show.legend = FALSE
  ) + 
  facet_wrap(~park_name, scales = "free") + #creates a different chart for each park_name and gives each chart their own scale (the default scale is fixed which means all charts would be on the same scale)
  theme_bw()

```

### Create these graphs as a base layer

```{r}

my_np_graph <- ggplot(ca_np, aes(year, visitors, group = park_name)) + 
  geom_point(
    aes(color = park_name,
        size = visitors),
    show.legend = FALSE
  ) + 
  facet_wrap(~park_name, scales = "free") + #creates a different chart for each park_name and gives each chart their own scale (the default scale is fixed which means all charts would be on the same scale)
  theme_bw()

```


### Export my graph as a .png

```{r}

ggsave(plot = my_np_graph, here("graphs", "np_graph.png"))

```







